---
title: "STA442 Homework4"
author: "Hanyu Zhou"
date: " 1003110858"
output:
  pdf_document: default
  word_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
# install.packages("INLA", repos=c(getOption("repos"), INLA="https://inla.r-inla-download.org/R/testing"), dep=TRUE)
# install.packages("Pmisc", repos = "http://r-forge.r-project.org")

library(mapmisc)
library(magrittr)    
library(dplyr)         
library(rpart)
library(partykit)
library(knitr)
library(tidyverse)
library(faraway)
library(lme4)      
library(SMPracticals)  
library(nlme)
library(INLA)
library(Pmisc)
library(Hmisc)
library(MEMSS)
library(MASS)
library(ggplot2)
library(kableExtra)
library(expm)
library('R.utils')
```

# Smoking

```{r ,warning=FALSE, message=FALSE, echo=FALSE}

### Model Code ###

smokeFile = Pmisc::downloadIfOld("http://pbrown.ca/teaching/appliedstats/data/smoke.RData")

load(smokeFile)
smoke = smoke[smoke$Age > 9, ]
forInla = smoke[, c("Age", "Age_first_tried_cigt_smkg",
"Sex", "Race", "state", "school", "RuralUrban")]
forInla = na.omit(forInla)
forInla$school = factor(forInla$school)

forSurv = data.frame(time = (pmin(forInla$Age_first_tried_cigt_smkg,
forInla$Age) - 4)/10, event = forInla$Age_first_tried_cigt_smkg <=
forInla$Age)
# left censoring
forSurv[forInla$Age_first_tried_cigt_smkg == 8, "event"] = 2
smokeResponse = inla.surv(forSurv$time, forSurv$event)

inla_formula = smokeResponse ~ Race + Sex + RuralUrban +
  f(school, model = "iid", hyper = list(prec = list(prior = "pc.prec", param = c(1.4,0.01)))) +
  f(state, model = "iid", hyper = list(prec = list(prior = "pc.prec", param = c(0.7,0.01))))


prof_model = inla(inla_formula,
                  control.family = list(variant = 1, hyper = list(alpha = list(prior = "normal", param = c(1,0.1)))),
                  control.mode = list(theta = c(8, 2, 5), restart = TRUE),
                  data = forInla, family = "weibullsurv", verbose = TRUE,
                  control.compute=list(config = TRUE))

```

## Introduction

The age at which children first try cigarette smoking is known to be earlier for males than females, earlier in rural areas than urban areas, and to vary by ethnicity. It is likely that significant variation amongst the US states exists, and that there is variation from one school to the next.

Concerning the 2014 American National Youth Tobacco Survey and the dataset on the pbrown.ca/teaching/appliedstats/data, I intended to investigate the following hypotheses:

1. Geographic variation (between states) in the mean age children first try cigarettes is substantially greater than variation amongst schools. As a result, tobacco control programs should target the states with the earliest smoking ages and not concern themselves with finding particular schools where smoking is a problem.
2. First cigarette smoking has a flat hazard function, or in other words is a first order Markov process. This means two non-smoking children have the same probability of trying cigarettes within the next month, irrespective of their ages but provided the known confounders (sex, rural/urban, etnicity) and random effects (school and state) are identical.

## Method

We need use survival analysis for this study since we intended to estimate the expected duration of time for children start smoking for the first time. And according to the shape of the data, I decided to use Weibull distribution to model it.

$$\begin{split}
Y_{ijk} &\sim \textrm{Weibull}(\rho_{ijk},\kappa)\\
\rho_{ijk} &= \exp(-\eta_{ijk})\\
\eta_{ijk} &= X_{ijk}\beta + U_{i} + V_{ij}\\
U_{i} &\sim N(0,\sigma_U^2)\\
V_{ij} &\sim N(0,\sigma_V^2)\\
\end{split}$$

where:

- $X_{ij}\beta$ is the subjects gender, ethnicity, whether they are from a rural or urban school
- $U_i$ is the school random effect.
- $V_{ij}$ is the state random effect.
- $\kappa$ is the Weibull shape parameter.

From the prior information: $exp(U_i) = 2 or 3$ but unlikely to see at 10, I know the range of $exp(U_i)$ is <=7, so $\sigma_U$ <= 1.7, and $\sigma_V$ is less than the half of the $\sigma_U$\
Therefore, I set the prior distributions as follows:
$$\begin{split}
\kappa \sim N(1,\ \ &0.1) \\
P(\sigma_U > 1.4) &= 0.01 \\
P(\sigma_V > 0.7) &= 0.01 \\
\end{split}$$

\newpage

## Result

```{r ,warning=FALSE, message=FALSE, echo=FALSE}
### model para ###

post.dat=round(exp(prof_model$mode$theta),2)
table1 <- 1/ sqrt(post.dat[2:3])
table1<-round(table1,4)
table1<-matrix(table1,nrow=1)

colnames(table1)<-c("SD of School", "SD of state")

knitr::kable(table1, caption="Posterior estimates")
```

From the following 4 plots, it's clear that:\
1, Geographic variation between states in the mean age children first try cigarettes is less than variation amongst schools. So tobacco control programs should actually target finding particular schools where smoking is a problem.\
2, According to the plot of prior distribution and the plot of hazard fuction, we can conclude the first cigarette smoking does not have a flat hazard function. And the non-smoking children with higher age have the higher probability of trying cigarettes within the next month provided the known confounders and random effects are identical.\


\newpage

```{r ,warning=FALSE, message=FALSE, echo=FALSE}
### porio plot

prof_model$priorPost = Pmisc::priorPost(prof_model)
for (Dparam in prof_model$priorPost$parameters) {
do.call(matplot, prof_model$priorPost[[Dparam]]$matplot)
}
do.call(legend, prof_model$priorPost$legend)
```

```{r ,warning=FALSE, message=FALSE, echo=FALSE}

forSurv$one = 1
hazEst = survfit(Surv(time, one) ~ 1, data=forSurv)
plot(hazEst, fun='cumhaz')

```




\newpage

# Death on the roads

```{r, warning=FALSE, echo=FALSE}

pedestrainFile = Pmisc::downloadIfOld("http://pbrown.ca/teaching/appliedstats/data/pedestrians.rds")

pedestrians = readRDS(pedestrainFile)
pedestrians = pedestrians[!is.na(pedestrians$time),]

pedestrians$y = pedestrians$Casualty_Severity == "Fatal"
pedestrians$timeCat = format(pedestrians$time,"%Y_%b_%a_h%H")

pedestrians$strata = paste(pedestrians$Light_Conditions,
                           pedestrians$Weather_Conditions,
                           pedestrians$timeCat)

theTable = table(pedestrians$strata, pedestrians$y)
onlyOne = rownames(theTable)[which(theTable[, 1] == 0 | theTable[, 2] == 0)]

x = pedestrians[!pedestrians$strata %in% onlyOne,]

theTable = table(pedestrians$strata, pedestrians$y)

onlyOne = rownames(theTable)[which(theTable[, 1] == 0 | theTable[, 2] == 0)]

theClogit = clogit(y ~ age + age:sex + strata(strata), data = x)

```

## Introduction

I used the data from www.gov.uk/government/statistical-data-sets/ras30-reportedcasualties-in-road-accidents, with all of the road traffic accidents in the UK from 1979 to 2015. The data below consist of all pedestrians involved in motor vehicle accidents with either fatal or slight injuries (pedestrians with moderate injuries have been removed).

According to the hypothesis, we investigated that whether men are involved in accidents more than women, and the proportion of accidents which are fatal is higher for men than for women. This might be due in part to women being more reluctant than men to walk outdoors late at night or in poor weather, and could also reflect men being on average more likely to engage in risky behaviour than women.\



## Method

First I fit a logistic regression model to examine the importance of chossing the lighting, weather and time conditions as strata.\

Then I used conditional logistic regression to model the data.\
We want\
$$\begin{split}
pr(Y_i &= 1|X_i) = \lambda_i\\
\log[\lambda_i/(1-\lambda_i)] &= \beta_0 + \sum_{p = 1}^{P} X_{ip} \beta_p\\
\end{split}$$
We have\
$$\begin{split}
pr(Y_i = 1|&X_i,Z_i = 1)   = \lambda_i^{*}\\
\log[\lambda_i^{*}/(1-\lambda_i^{*})]&= \beta_0^{*} + \sum_{p = 1}^{P} X_{ip} \beta_p^{*}\\
\end{split}$$
By the theorem we get:\
$$\begin{split}
\beta_p^{*} = \beta_0 + log[{pr(Z_i = 1|Y_i = 1)}/{pr(Z_i=1|Y_i=0)}] \textrm{  if} \ \ p = 0 \\
\beta_p^{*} = \beta_p \textrm{  if} \ \ p \ne 0 \\
\end{split}$$
where:

- $X_{ip}\beta$ is the subjects' gender, and age.
- $Y_{i}$ is the status of casualty.
- $Z_{i}$ is the strata of lightness, weather and time conditions.

\newpage

## Result

```{r, warning=FALSE, echo=FALSE}

model2 = glm(y ~ sex + age + Light_Conditions + Weather_Conditions, data = x, family = "binomial")
summary(model2)$coef[13:22, ]

theCoef = rbind(as.data.frame(summary(theClogit)$coef), `age 26 - 35` = c(0, 1, 0, NA, NA))
theCoef$sex = c("Male", "Female")[1 + grepl("Female", rownames(theCoef))]
theCoef$age = as.numeric(gsub("age|Over| - [[:digit:]].*|[:].*", "", rownames(theCoef)))
theCoef = theCoef[order(theCoef$sex, theCoef$age), ]

knitr::kable(theCoef, caption="The coeficients of conditional logistic regression")
```

\newpage
1,The summary of coefficients of logistics regression model is listed as above, we can see the p-value of lighting and weather conditions are extremely small, it means the lighting and weather conditions are both statistically significant, so they have notable influence to the outcome. And the time condition have relationship with lighting condition(darkness at night). Therefore I set them as strata.\
2,The coefficients of conditional logistic regression are summarized in the table. The reference group is the male with age from 26 to 35(coef = 0). It is clear that men are more likely to involved in accidents than women in the rough. After age 35, the proportion of accidents which are fatal is higher for men than for women, but the proportion is pretty much the same from age 0 to 35. Also it's notable that wamen tend to be safer particularly as middle-aged(26-45), rather than as teenagers and in early adulthood.


\newpage

# Appendix
```{r ,warning=FALSE, message=FALSE, echo=TRUE}

# Smoking

### Model Code ###

smokeFile = Pmisc::downloadIfOld("http://pbrown.ca/teaching/appliedstats/data/smoke.RData")

load(smokeFile)
smoke = smoke[smoke$Age > 9, ]
forInla = smoke[, c("Age", "Age_first_tried_cigt_smkg",
"Sex", "Race", "state", "school", "RuralUrban")]
forInla = na.omit(forInla)
forInla$school = factor(forInla$school)

forSurv = data.frame(time = (pmin(forInla$Age_first_tried_cigt_smkg,
forInla$Age) - 4)/10, event = forInla$Age_first_tried_cigt_smkg <=
forInla$Age)
# left censoring
forSurv[forInla$Age_first_tried_cigt_smkg == 8, "event"] = 2
smokeResponse = inla.surv(forSurv$time, forSurv$event)

inla_formula = smokeResponse ~ Race + Sex + RuralUrban +
  f(school, model = "iid", hyper = list(prec = list(prior = "pc.prec", param = c(1.3,0.01)))) +
  f(state, model = "iid", hyper = list(prec = list(prior = "pc.prec", param = c(0.6,0.01))))


prof_model = inla(inla_formula,
                  control.family = list(variant = 1, hyper = list(alpha = list(prior = "normal", param = c(1,0.1)))),
                  control.mode = list(theta = c(8, 2, 5), restart = TRUE),
                  data = forInla, family = "weibullsurv", verbose = TRUE,
                  control.compute=list(config = TRUE))


### model para ###

post.dat=round(exp(prof_model$mode$theta),2)
table1 <- 1/ sqrt(post.dat[2:3])
table1<-round(table1,4)
table1<-matrix(table1,nrow=1)

colnames(table1)<-c("SD of School", "SD of state")

knitr::kable(table1, caption="Posterior estimates")

### porio plot

prof_model$priorPost = Pmisc::priorPost(prof_model)
for (Dparam in prof_model$priorPost$parameters) {
do.call(matplot, prof_model$priorPost[[Dparam]]$matplot)
}
do.call(legend, prof_model$priorPost$legend)

forSurv$one = 1
hazEst = survfit(Surv(time, one) ~ 1, data=forSurv)
plot(hazEst, fun='cumhaz')


# Death on the roads
pedestrainFile = Pmisc::downloadIfOld("http://pbrown.ca/teaching/appliedstats/data/pedestrians.rds")

pedestrians = readRDS(pedestrainFile)
pedestrians = pedestrians[!is.na(pedestrians$time),]

pedestrians$y = pedestrians$Casualty_Severity == "Fatal"
pedestrians$timeCat = format(pedestrians$time,"%Y_%b_%a_h%H")

pedestrians$strata = paste(pedestrians$Light_Conditions,
                           pedestrians$Weather_Conditions,
                           pedestrians$timeCat)

theTable = table(pedestrians$strata, pedestrians$y)
onlyOne = rownames(theTable)[which(theTable[, 1] == 0 | theTable[, 2] == 0)]

x = pedestrians[!pedestrians$strata %in% onlyOne,]

theTable = table(pedestrians$strata, pedestrians$y)

onlyOne = rownames(theTable)[which(theTable[, 1] == 0 | theTable[, 2] == 0)]
theClogit = clogit(y ~ age + age:sex + strata(strata), data = x)
model2 = glm(y ~ sex + age + Light_Conditions + Weather_Conditions, data = x, family = "binomial")
summary(model2)$coef
theCoef = rbind(as.data.frame(summary(theClogit)$coef), `age 26 - 35` = c(0, 1, 0, NA, NA))
theCoef$sex = c("Male", "Female")[1 + grepl("Female", rownames(theCoef))]
theCoef$age = as.numeric(gsub("age|Over| - [[:digit:]].*|[:].*", "", rownames(theCoef)))
theCoef = theCoef[order(theCoef$sex, theCoef$age), ]

knitr::kable(theCoef, caption="The coeficients of conditional logistic regression")

```


